Methods for Large Scale Learning to Rank

### 2014-03-17 report/Chapters/BenchmarkResults.tex
begin LETOR 4.0 results

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment, I will only look at the standard Learning-to-Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
Queries & 19,994 & 2,994 & 6,983 \\ 
Documents & 473,134 & 71,083 & 165,660 \\ 
Features & 519 & 519 & 519 \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper \cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

@@

\section{LETOR}
The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time. LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009. The LETOR3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\

\subsection{LETOR3.0}
The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991. The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004. Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\

Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{Craswell2003}:
\begin{description}
\item[topic distillation (TD)]these queries involve finding relevant pages given a broad query. E.g., the query 'cotton industry' which is entered with the aim to find pages that provide information about the cotton industry.
\item[named page finding (NP)]in these queries the user asks for one specific page by name. E.g., the user queries for 'ADA Enforcement' to find the page $http://www.usdoj.gov/crt/ada/enforce.htm$.
\item[home page finding (HP)]these queries are similar to NP queries in that the user is looking for one specific page, but now this specific page is a homepage. E.g., the user queries for 'Tennessee Valley Authority' to find the homepage $http://www.tva.gov$.
\end{description}
With query sets available from both the TREC 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\

@@

I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline. We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value. Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet. Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
\begin{table}
\begin{tabular}{l|ll}
 & \ac{nDCG}@10 & \ac{MAP} \\ 
 \hline
TD2003 & 0.348 & 0.2753 \\ 
TD2004 & 0.317 & 0.2231 \\ 
NP2003 & 0.801 & 0.6895 \\ 
NP2004 & 0.812 & 0.6720 \\ 
HP2003 & 0.837 & 0.7659 \\ 
HP2004 & 0.784 & 0.6899 \\ 
OHSUMED & 0.441 & 0.4457 \\ 
\end{tabular}
\caption{Performance of ListNet on LETOR}
\label{tbl:LETOR_ListNet}
\end{table}

\begin{table}
\begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
 & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{RankSVM-Primal} &\rotatebox{55}{RankSVM-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
 \hline
TD2003 & 0.3297 & 0.3571 & 0.3467 & 0.3367 & 0.348 \\ 
TD2004 & 0.2832 & 0.2913 & 0.3090 & 0.3343 & 0.317 \\ 
NP2003 & 0.8025 & 0.7894 & 0.7955 & 0.7986 & 0.801 \\ 
NP2004 & 0.8040 & 0.7950 & 0.7977 & 0.8075 & 0.812 \\ 
HP2003 & 0.8216 & 0.8180 & 0.8162 & 0.8325 & 0.837 \\ 
HP2004 & 0.7188 & 0.7720 & 0.7666 & 0.8221 & 0.784 \\ 
OHSUMED & 0.4436 & 0.4504 & 0.4523 & 0.4568 & 0.441 \\ 
\# winning datasets & 2 & 2 & 1 & 3 & - \\ 
\end{tabular}
\caption{Comparison of algorithms recently evaluated on LETOR3.0 with the ListNet baselines}
\label{tbl:LETOR_recently_added}
\end{table}
\subsection{LETOR4.0}
\subsubsection{Results}
BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.

>>>>

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment, I will only look at the standard Learning-to-Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
\# of queries & 19,994 & 2,994 & 6,983 \\ 
\# of documents & 473,134 & 71,083 & 165,660 \\ 
\# of features & 519 & 519 & 519 \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper \cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

@@

\section{LETOR}
The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time. LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.\\

\subsection{LETOR 3.0}
The LETOR 3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection. The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991. The .gov collection is a web crawl obtained in January 2002, which was used for the \ac{TREC} web track in 2003 and 2004. Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\

Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{Craswell2003}:
\begin{description}
\item[topic distillation (TD)]these queries involve finding relevant pages given a broad query. E.g., the query 'cotton industry' which is entered with the aim to find pages that provide information about the cotton industry.
\item[named page finding (NP)]in these queries the user asks for one specific page by name. E.g., the user queries for 'ADA Enforcement' to find the page $http://www.usdoj.gov/crt/ada/enforce.htm$.
\item[home page finding (HP)]these queries are similar to NP queries in that the user is looking for one specific page, but now this specific page is a homepage. E.g., the user queries for 'Tennessee Valley Authority' to find the homepage $http://www.tva.gov$.
\end{description}
With query sets available from both the \ac{TREC} 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\

@@

I will describe the performance of the algorithms that were evaluated on LETOR 3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline. We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value. Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet. Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
\begin{table}
\begin{tabular}{l|ll}
 & \ac{nDCG}@10 & \ac{MAP} \\ 
 \hline
TD2003 & 0.348 & 0.2753 \\ 
TD2004 & 0.317 & 0.2231 \\ 
NP2003 & 0.801 & 0.6895 \\ 
NP2004 & 0.812 & 0.6720 \\ 
HP2003 & 0.837 & 0.7659 \\ 
HP2004 & 0.784 & 0.6899 \\ 
OHSUMED & 0.441 & 0.4457 \\ 
\end{tabular}
\caption{Performance of ListNet on LETOR 3.0}
\label{tbl:LETOR_ListNet}
\end{table}

\begin{table}
\begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
 & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{RankSVM-Primal} &\rotatebox{55}{RankSVM-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
 \hline
TD2003 & 0.3297 & 0.3571 & 0.3467 & 0.3367 & 0.348 \\ 
TD2004 & 0.2832 & 0.2913 & 0.3090 & 0.3343 & 0.317 \\ 
NP2003 & 0.8025 & 0.7894 & 0.7955 & 0.7986 & 0.801 \\ 
NP2004 & 0.8040 & 0.7950 & 0.7977 & 0.8075 & 0.812 \\ 
HP2003 & 0.8216 & 0.8180 & 0.8162 & 0.8325 & 0.837 \\ 
HP2004 & 0.7188 & 0.7720 & 0.7666 & 0.8221 & 0.784 \\ 
OHSUMED & 0.4436 & 0.4504 & 0.4523 & 0.4568 & 0.441 \\ 
\# winning datasets & 2 & 2 & 1 & 3 & - \\ 
\end{tabular}
\caption{\ac{nDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
\label{tbl:LETOR_recently_added}
\end{table}
\subsection{LETOR 4.0}
The LETOR 4.0 benchmark collection\footnote{http://http://research.microsoft.com/en-us/um/beijing/projects/letor/} consists of the Gov-2 document collection and two query datasets from the Million Query Track at \ac{TREC} 2007 (MQ2007) and \ac{TREC} 2008 (MQ2008). LETOR 4.0 consists of a semi-supervised ranking task, a rank aggregation task and a listwise ranking task next to the supervised ranking task. Table \ref{tab:letor4_characteristics} shows the collection characteristics in number of queries, documents and features for LETOR 4.0. Evaluation on this collection is performed using a five-fold cross-validation, where partitioning of the data into folds was performed beforehand by the creators of the MQ2007 and MQ2008 datasets. Documents in the dataset were

\begin{table}
\begin{tabular}{l|ll}
 Dataset & MQ2007 & MQ2008 \\ 
 \hline
 Queries & 1692 & 784 \\ 
 Documents & 69,622 & 15,211 \\
 Features & 46 & 46 \\
\end{tabular}
\label{tab:letor4_characteristics}
\caption{Characteristics of the LETOR 4.0 collection}
\end{table}
\subsubsection{Results}
RankSVM-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 dataset and were implemented and evaluated by the publishers of LETOR 4.0. Table \ref{tab:letor4_baseline_results} shows the performance of those baseline models on the LETOR 4.0 benchmark collection.\\

\begin{table}
\begin{tabular}{l|lll}
Model & Dataset & Mean \ac{nDCG} & \ac{nDCG}@10 \\ 
\hline
RankSVM-Struct & MQ2007 & 0.4966 & 0.4439 \\ 
 & MQ2008 & 0.4832 & 0.2279 \\ 
\hline
ListNet & MQ2007 & 0.4988 & 0.4440 \\ 
 & MQ2008 & 0.4914 & 0.2303 \\ 
\hline
AdaRank-\ac{nDCG} & MQ2007 & 0.4914 & 0.4369 \\ 
 & MQ2008 & 0.4950 & 0.2307 \\ 
\hline
AdaRank-\ac{MAP} & MQ2007 & 0.4891 & 0.4335 \\ 
 & MQ2008 & 0.4915 & 0.2288 \\ 
\hline
RankBoost & MQ2007 & 0.5003 & 0.4464 \\ 
 & MQ2008 & 0.4850 & 0.2255 \\ 
\end{tabular}
\caption{Comparison of LETOR 4.0 baseline models}
\label{tab:letor4_baseline_results}
\end{table}

BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.
